\vspace{-10pt}Predictive Quantization for MIMO-OFDM SVD Precoders using Reservoir Computing Framework \vspace{-10pt}

### 2020-01-04 paper/main.tex
Spelling fix
Adding minor changes (removing not necesarry citations etc)

Combining these two inherent properties (viz. manifold structure and correlations) to frame manifold based predictive quantization algorithms requires a generalization of various well known linear algorithms for prediction, quantization and interpolation, to the underlying non-linear manifold structure. This is typically done by exploiting the differential geometry of the manifold, and relevant past work in this direction has been presented in \cite{Gupt1905:Predictive,6891198,6545375,5671092,Li2016,krishnamachari2013geometry,chang2011adaptive}. Manifold based approaches for predictive quantization have provided significant performance benefits, given that they work with the lower number of effective dimensions while framing the manifold operations to predict and quantize.
However, these approaches are non-trivial, since they require specific operations over the non-linear manifold differential geometry.
In this paper, we aim to capture the above non-linear relations using a reservoir computing based predictive quantizer to exploit the temporal correlations effectively.
The proposed framework offers a simple solution to the predictive quantization problem, which brings about ease of implementation, along with improved performance, when compared to \cite{6891198}, which proposed a corresponding manifold based method to the problem considered in this paper.

@@

The input to the reservoir $i$ that corresponds to subcarrier $i$ is $\tilde{\bu}_{i,t-1}$ (vectorized representation of $\tilde{\bU}_{i,t-1}$), with dimensions $D_{\text{in}}=2N_TN_R-N_R^2+N_R$.
Using a randomly initialized $D_{\text{resv}}\times D_{\text{in}}$ input coupler matrix $\bW^{\text{in}}_i$, the $D_{\text{in}}$ dimensional input is mapped to a $D_{\text{resv}}(>>D_{\text{in}})$ dimensional vector, where $D_{\text{resv}}$ is the dimension of the reservoir state vector $\br_{i,t}$. $\br_{i,t}$ is initialized with $D_{\text{resv}}$ zeros, and is updated via the following equation:
\begin{equation}
\br_{i,t}=\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})
\label{res_upd}
\end{equation}
where $\bA_i$ is the adjacency matrix that captures the reservoir dynamics, and $\text{tanh}$ is applied element wise.
Typical choices for $\bA_i$ have been Erdos-Renyi graphs with an upper bounded maximum eigenvalue \cite{mosleh2017brain,pathak2017using}. The output of the reservoir is $\bp_{i,t}$ (vectorized representation of $\bP_{i,t}$), which denotes the predicted precoder at $i$-th subcarrier, $t$-th time instant.
Therefore, the output of the reservoir is also a $D_{\text{in}}$ dimensional vector (viz. $D_{\text{out}}=D_{\text{in}}$).
The output $\bp_{i,t}$ is obtained using a matrix transformation of the reservoir state $\br_{i,t}$ with $D_{\text{out}}\times D_{\text{resv}}$ output coupler matrix $\bW^{\text{out}}_{i,t-1}$, which forms the prediction function $f_P(\cdot)$ with $\tilde{\bU}_{i,t-1}$ as its argument, for the RC framework
\begin{equation}
\bp_{i,t}=\bW^{\text{out}}_{i,t-1}\br_{i,t}=f_P(\tilde{\bU}_{i,t-1})
\label{pred_op}
\end{equation}

@@

We now describe the quantization function $q_P(\cdot)$ which is essentially the same as considered in \cite{Gupt1905:Predictive,6891198}, that exploits the fact that tangent spaces local to a point in manifold are vector spaces. To do so, the quantization algorithm quantizes the local tangent space at the predicted precoder $\bP_{i,t}$ to get a codebook, $T^\mathcal{C}$ having $2^{\mathcal{B}}$ codeword tangents in the tangent space. The quantized precoders are obtained by traversing along the `codeword tangents' from the predicted precoder $\bP_{i,t}$, using a retraction map (denoted by $\text{retract}(\cdot)$). The receiver conveys the feedback $\text{fb}_{i,t}$ to the transmitter, which minimizes the chordal distance metric ($d_s(\bX,\bY)$ for $\bX,\bY \in \text{St}(N_T,N_R)$ detailed in \cite{Gupt1905:Predictive,6891198}) between the quantized precoders and the ideal precoder $\bU_{i,t}$ estimated from SVD of $\bH_{i,t}$ \eqref{eqn:cn1}
 \begin{equation}
\label{eqn:cn1}
\text{fb}_{i,t} \gets \text{argmin}_{i \in \{1,2,\ldots,2^{\mathcal{B}}\}}\left (d_s\left (\bU_{i,t},\text{retract}(\bP_{i,t},T^\mathcal{C}[i])\right ) \right)
\end{equation}
The transmitter utilizes the above quantization function $q_P$ to obtain $\tilde{\bU}_{i,t}=q_P(\bp_{i,t},\text{fb}_{i,t})$ from $\bp_{i,t},\text{fb}_{i,t}$.
\begin{equation}
\tilde{\bU}_{i,t}=q_P(\bP_{i,t},\text{fb}_{i,t})=\text{retract}(\bP_{i,t},T^\mathcal{C}[\text{fb}_{i,t}])
\label{eqn:cn2}
\end{equation}
% Given two points $\bX$ and $\bY$ in $\text{St}(N_T,N_R)$, a lifting operation $\bT^Y_X=\texttt{lift}(\bX,\bY)$, $\texttt{lift}:\text{St}(N_T,N_R)\times \text{St}(N_T,N_R) \to \mathcal{T}_X \text{St}(N_T,N_R)$ yields a tangent from $X$ to $Y$, denoted as $\bT^Y_X \in \mathcal{T}_X \text{St}(N_T,N_R)$, where $\mathcal{T}_X \text{St}(N_T,N_R)$ is the local tangent space at $\bX$.
% A corresponding retraction operation $Y=\texttt{retract}(X,\bT^Y_X)$, $\texttt{retract}: \text{St}(N_T,N_R)\times \mathcal{T}_X \text{St}(N_T,N_R) \to \text{St}(N_T,N_R)$ gives back the manifold point obtained by traversing in the tangent direction given by the second argument ($\bT^Y_X$).
% In this work, the chosen lifting-retraction pairs are the Cayley exponentials, discussed in \cite{DBLP:journals/corr/abs-1708-00045,Gupt1905:Predictive}. % and used in \cite{Gupt1905:Predictive}.
% The Cayley exponential lifting operation maps two points in $\text{St}(N_T,N_R)$, to a $N_T\times N_T$ skew Hermitian matrix that represents the tangent from first point to the other.

@@

To evaluate the complete MIMO-OFDM setting, predictive quantization algorithms should be considered in conjunction with interpolation algorithms to estimate the precoders at non fed-back points.
We perform simulation for the complete MIMO-OFDM setting (visually illustrated in Fig. 2) with $N=64$ subcarriers, and receiver feeding back quantized precoder information for $8$ subcarriers, and transmitter utilizing Caley interpolation to estimate the precoder at other subcarriers. The transmitter obtains feed back for $8$ evenly spaced subcarriers indexed from 0, viz. $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$). For non fed back subcarriers $i\neq 9k$, the Cayley method \cite{Gupt1905:Predictive} is used to interpolate $\tilde{\bU}_{i,t}$, from available estimates, $\tilde{\bU}_{9n,t}$, $\tilde{\bU}_{9(n+1), t}$

>>>>

Combining these two inherent properties (viz. manifold structure and correlations) to frame manifold based predictive quantization algorithms requires a generalization of various well known linear algorithms for prediction, quantization and interpolation, to the underlying non-linear manifold structure. This is typically done by exploiting the differential geometry of the manifold, and relevant past work in this direction has been presented in \cite{Gupt1905:Predictive,6891198,6545375,Li2016,chang2011adaptive}. Manifold based approaches for predictive quantization have provided significant performance benefits, given that they work with the lower number of effective dimensions while framing the manifold operations to predict and quantize.
However, these approaches are non-trivial, since they require specific operations over the non-linear manifold differential geometry.
In this paper, we aim to capture the above non-linear relations using a reservoir computing based predictive quantizer to exploit the temporal correlations effectively.
The proposed framework offers a simple solution to the predictive quantization problem, which brings about ease of implementation, along with improved performance, when compared to \cite{6891198}, which proposed a corresponding manifold based method to the problem considered in this paper.

@@

The input to the reservoir $i$ that corresponds to subcarrier $i$ is $\tilde{\bu}_{i,t-1}$ (vectorized representation of $\tilde{\bU}_{i,t-1}$), with dimensions $D_{\text{in}}=2N_TN_R-N_R^2+N_R$.
Using a randomly initialized $D_{\text{resv}}\times D_{\text{in}}$ input coupler matrix $\bW^{\text{in}}_i$, the $D_{\text{in}}$ dimensional input is mapped to a $D_{\text{resv}}(>>D_{\text{in}})$ dimensional vector, where $D_{\text{resv}}$ is the dimension of the reservoir state vector $\br_{i,t}$. $\br_{i,t}$ is initialized with $D_{\text{resv}}$ zeros, and is updated via the following equation:
\begin{equation}
\br_{i,t}=\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})
\label{res_upd}
\end{equation}
where $\bA_i$ is the adjacency matrix that captures the reservoir dynamics, and $\text{tanh}$ is applied element wise.
Typical choices for $\bA_i$ have been Erdos-Reny\"i graphs with an upper bounded maximum eigenvalue \cite{mosleh2017brain,pathak2017using}. The output of the reservoir is $\bp_{i,t}$ (vectorized representation of $\bP_{i,t}$), which denotes the predicted precoder at $i$-th subcarrier, $t$-th time instant.
Therefore, the output of the reservoir is also a $D_{\text{in}}$ dimensional vector (viz. $D_{\text{out}}=D_{\text{in}}$).
The output $\bp_{i,t}$ is obtained using a matrix transformation of the reservoir state $\br_{i,t}$ with $D_{\text{out}}\times D_{\text{resv}}$ output coupler matrix $\bW^{\text{out}}_{i,t-1}$, which forms the prediction function $f_P(\cdot)$ with $\tilde{\bU}_{i,t-1}$ as its argument, for the RC framework
\begin{equation}
\bp_{i,t}=\bW^{\text{out}}_{i,t-1}\br_{i,t}=f_P(\tilde{\bU}_{i,t-1})
\label{pred_op}
\end{equation}

@@

We now describe the quantization function $q_P(\cdot)$ which is essentially the same as considered in \cite{Gupt1905:Predictive,6891198}, that exploits the fact that tangent spaces local to a point in manifold are vector spaces. To do so, the quantization algorithm quantizes the local tangent space at the predicted precoder $\bP_{i,t}$ to get a codebook, $T^\mathcal{C}$ having $2^{\mathcal{B}}$ codeword tangents in the tangent space. The quantized precoders are obtained by traversing along the `codeword tangents' from the predicted precoder $\bP_{i,t}$, using a retraction map (denoted by $\text{retract}(\cdot)$). The receiver conveys the feedback $\text{fb}_{i,t}$ to the transmitter, which minimizes the chordal distance metric ($d_s(\bX,\bY)$ for $\bX,\bY \in \text{St}(N_T,N_R)$ \cite{Gupt1905:Predictive,6891198}) between the quantized precoders and the ideal precoder $\bU_{i,t}$ estimated from SVD of $\bH_{i,t}$ \eqref{eqn:cn1}
 \begin{equation}
\label{eqn:cn1}
\text{fb}_{i,t} \gets \text{argmin}_{i \in \{1,2,\ldots,2^{\mathcal{B}}\}}\left (d_s\left (\bU_{i,t},\text{retract}(\bP_{i,t},T^\mathcal{C}[i])\right ) \right)
\end{equation}
The transmitter utilizes the above quantization function $q_P$ to obtain $\tilde{\bU}_{i,t}=q_P(\bp_{i,t},\text{fb}_{i,t})$ from $\bp_{i,t},\text{fb}_{i,t}$.
\begin{equation}
\tilde{\bU}_{i,t}=q_P(\bP_{i,t},\text{fb}_{i,t})=\text{retract}(\bP_{i,t},T^\mathcal{C}[\text{fb}_{i,t}])
\label{eqn:cn2}
\end{equation}
% Given two points $\bX$ and $\bY$ in $\text{St}(N_T,N_R)$, a lifting operation $\bT^Y_X=\texttt{lift}(\bX,\bY)$, $\texttt{lift}:\text{St}(N_T,N_R)\times \text{St}(N_T,N_R) \to \mathcal{T}_X \text{St}(N_T,N_R)$ yields a tangent from $X$ to $Y$, denoted as $\bT^Y_X \in \mathcal{T}_X \text{St}(N_T,N_R)$, where $\mathcal{T}_X \text{St}(N_T,N_R)$ is the local tangent space at $\bX$.
% A corresponding retraction operation $Y=\texttt{retract}(X,\bT^Y_X)$, $\texttt{retract}: \text{St}(N_T,N_R)\times \mathcal{T}_X \text{St}(N_T,N_R) \to \text{St}(N_T,N_R)$ gives back the manifold point obtained by traversing in the tangent direction given by the second argument ($\bT^Y_X$).
% In this work, the chosen lifting-retraction pairs are the Cayley exponentials, discussed in \cite{DBLP:journals/corr/abs-1708-00045,Gupt1905:Predictive}. % and used in \cite{Gupt1905:Predictive}.
% The Cayley exponential lifting operation maps two points in $\text{St}(N_T,N_R)$, to a $N_T\times N_T$ skew Hermitian matrix that represents the tangent from first point to the other.

@@

To evaluate the complete MIMO-OFDM setting, predictive quantization algorithms should be considered in conjunction with interpolation algorithms to estimate the precoders at non fed-back points.
We perform simulation for the complete MIMO-OFDM setting (visually illustrated in Fig. 2) with $N=64$ subcarriers. The transmitter obtains feed back for $8$ evenly spaced subcarriers indexed from 0, viz. $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$). For non fed back subcarriers $i\neq 9k$, the Cayley method \cite{Gupt1905:Predictive} is used to interpolate $\tilde{\bU}_{i,t}$, from available estimates, $\tilde{\bU}_{9n,t}$, $\tilde{\bU}_{9(n+1), t}$
